Monitoring misinformation related interventions by Facebook, Twitter and YouTube: methods and illustration.

### 2022-05-09 article/draft.tex
final typose

Hence, in the present article, we focus on mainstream platforms' policies and interventions regarding content with low credibility or false information, commonly referred to as {\it Fake News} (see Lazer et al. (2018)~\cite{lazer}). The misinformation phenomenon is still ill-defined by the academic community, as it encompasses several combined features such as spreading inaccurate, false or misleading information, with or without the intention of influencing or manipulating a target pool of audience.  The growth of social networking platforms over the last decade in terms of number of users worldwide and volume of content, has modified the information ecosystem in terms of production of information and its mediation. Many users can now produce and share content which includes news related information, without having to abide by strict editorial processes that ensure accuracy of information and reliability of sources. In particular, false or inaccurate content produced and shared on social networking platforms concerning the political life or public health may have a potentially harmful impact on the society when it goes viral. This gave rise to a set of heterogenous policies across mainstream platforms, mainly based on fact-checking but also content moderators and users' reporting. For example, Facebook has a substantial partnership program with Fact-checking partners certified by the non-partisan International Fact-Checking Network. The company uses a number of signals and machine learning models to predict misinformation and surface it to fact-checkers.\footnote{See the \href{https://www.facebook.com/journalismproject/programs/third-party-fact-checking/how-it-works}{section Frequently asked questions: `How does Facebook use technology to detect potential misinformation?''}} Twitter seems to have a different approach where they focus on providing context rather than fact-checking\footnote{To the best of our knowledge, Twitter does not have a page which summarizes its fact-checking strategy. The Twitter Safety Team tweeted on June 3, 2020 the following:  ``We heard: 1. Twitter shouldn’t determine the truthfulness of Tweets 2. Twitter should provide context to help people make up their own minds in cases where the substance of a Tweet is disputed. Hence, our focus is on providing context, not fact-checking.'' Tweet ID \href{https://twitter.com/TwitterSafety/status/1267986503721988096}{1267986503721988096}.} and the platform is testing a new system based on the wisdom of the crowds to tackle misinformation (see \href{https://blog.twitter.com/en\_us/topics/product/2021/introducing-birdwatch-a-community-based-approach-to-misinformation}{Twitter Birdwatch}). As for YouTube, this platform utilizes the \href{https://schema.org/ClaimReview}{schema.org ClaimReview} markup, where fact-checking articles created by eligible publishers can appear on information panels (see Appendix \ref{links} for references).

@@

\begin{figure}[h]
	\centering
		\includegraphics[scale=0.21]{./img/oann/fig3_oann.png}
		\includegraphics[scale=0.31]{./img/tony/fig3_tony.png}
	\caption{Left Panel: Tweet announcing OANN moving to Rumble, Twitter ID \href{https://twitter.com/OANN/status/1372238828425998336}{1372238828425998336}. Right Panel: Tony Heller's tweet after getting suspended from YouTube, Twitter ID \href{https://twitter.com/Tony\_Heller/status/1310703852769796097}{1310703852769796097}.}
	\label{tweets_about_yt_suspensions}
\end{figure}

Tony Heller writes regularly blog posts on the website {\it realclimatescience.com}, which also has a ``low'' factual reporting score according to the website Media Bias Fact Check.\footnote{See \href{https://mediabiasfactcheck.com/real-climate-science/}{mediabiasfactcheck.com/real-climate-science/}}
Tony Heller tweeted on September 29, 2020 about his YouTube channel being `shut down' because of a video about an anti-covid-lockdown doctor getting arrested (see the right panel of Figure \ref{tweets_about_yt_suspensions}).
We applied the same methods as in the previous example to collect data.
Again a one-week suspension period from September $29$ until October $5$ can be observed clearly in the daily number of uploaded videos and of views (see Figure \ref{delete_youtube_tony}).
Comparing one month before and after the suspension, this channel also witnessed a drop of view counts by $-70\%$.

@@

As most YouTube channels upload videos on a weekly or even a monthly basis, a one-week suspension does not appear very restrictive at first sight. 
But we observed that the intervention was followed by a reduction in the number of view for the two channels investigated, with one channel even stopping its YouTube activity in the following months. 
Although more research is needed before drawing a definitive conclusion, it is possible that the one-week suspension following a strike might reduce the reach of the very active YouTube channels.

@@

%\begin{table}[h]
%\begin{tabular}{|c|c|c|c|}
%\hline
%\multicolumn{4}{|c|}{Number of posts}                                                                                                                    \\ \hline
%without a flag & with an information flag & with a fact-check flag & \begin{tabular}[c]{@{}c@{}}with a fact-check and an \\  information flags\end{tabular} \\ \hline
%0            & 7                        & 4                      & 7                                                                                     \\ \hline
%\end{tabular}
%\caption{Count for the Facebook posts with the different types of flags having shared a link fact-checked False by one of Facebook’s partners}
%\label{tab_flags_fb}
%\end{table}

@@

As we collected the recommendations of misinformation videos, we expected to find a high proportion of videos of questionable content.
But only {$18.4\%$} of the recommendations corresponded to $misleading$ or {\it hyper-partisan} videos.
This relatively small percentage hints towards that idea that YouTube is probably reducing the visibility of videos containing misinformation.
However, when we investigated misinformation at the channel level, we found that videos belonging to channels that are not highly authoritative were being often recommended.
It appears that the recommendation system of YouTube is reducing the visibility of certain videos but not of unreliable channels.

>>>>

Hence, in the present article, we focus on mainstream platforms' policies and interventions regarding content with low credibility or false information, commonly referred to as {\it Fake News} (see Lazer et al. (2018)~\cite{lazer}). The misinformation phenomenon is still ill-defined, as it encompasses several combined features such as spreading inaccurate, false or misleading information, with or without the intention of influencing or manipulating a target pool of audience.  The growth of social networking platforms over the last decade in terms of number of users worldwide and volume of content, has modified the information ecosystem in terms of production of information and its mediation. Many users can now produce and share content which includes news related information, without having to abide by strict editorial processes that ensure accuracy of information and reliability of sources. In particular, false or inaccurate content produced and shared on social networking platforms concerning the political life or public health may have a potentially harmful impact on the society when it goes viral. This gave rise to a set of heterogenous policies across mainstream platforms, mainly based on fact-checking but also content moderators and users' reporting. For example, Facebook has a substantial partnership program with Fact-checking partners certified by the non-partisan International Fact-Checking Network. The company uses a number of signals and machine learning models to predict misinformation and surface it to fact-checkers.\footnote{See the \href{https://www.facebook.com/journalismproject/programs/third-party-fact-checking/how-it-works}{section Frequently asked questions: `How does Facebook use technology to detect potential misinformation?''}} Twitter seems to have a different approach where they focus on providing context rather than fact-checking\footnote{To the best of our knowledge, Twitter does not have a page which summarizes its fact-checking strategy. The Twitter Safety Team tweeted on June 3, 2020 the following:  ``We heard: 1. Twitter shouldn’t determine the truthfulness of Tweets 2. Twitter should provide context to help people make up their own minds in cases where the substance of a Tweet is disputed. Hence, our focus is on providing context, not fact-checking.'' Tweet ID \href{https://twitter.com/TwitterSafety/status/1267986503721988096}{1267986503721988096}.} and the platform is testing a new system based on the wisdom of the crowds to tackle misinformation (see \href{https://blog.twitter.com/en\_us/topics/product/2021/introducing-birdwatch-a-community-based-approach-to-misinformation}{Twitter Birdwatch}). As for YouTube, this platform utilizes the \href{https://schema.org/ClaimReview}{schema.org ClaimReview} markup, where fact-checking articles created by eligible publishers can appear on information panels (see Appendix \ref{links} for references).

@@

\begin{figure}[h]
	\centering
		\includegraphics[scale=0.21]{./img/oann/fig3_oann.png}
		\includegraphics[scale=0.31]{./img/tony/fig3_tony.png}
	\caption{Left Panel: Tweet announcing OANN moving to Rumble, Twitter ID \href{https://twitter.com/OANN/status/1372238828425998336}{1372238828425998336}. Right Panel: Tony Heller's tweet after getting temporally suspended from YouTube.}
	\label{tweets_about_yt_suspensions}
\end{figure}
%Twitter ID \href{https://twitter.com/Tony\_Heller/status/1310703852769796097}{1310703852769796097}
\smallskip

Tony Heller writes regularly blog posts on the website {\it realclimatescience.com}, which also has a ``low'' factual reporting score according to the website Media Bias Fact Check.\footnote{See \href{https://mediabiasfactcheck.com/real-climate-science/}{mediabiasfactcheck.com/real-climate-science/}}
He tweeted on September 29, 2020 about his YouTube channel being `shut down' because of a video about an anti-covid-lockdown doctor getting arrested (see the right panel of Figure \ref{tweets_about_yt_suspensions}).
We applied the same methods as in the previous example to collect data.
Again a one-week suspension period from September $29$ until October $5$ can be observed clearly in the daily number of uploaded videos and of views (see Figure \ref{delete_youtube_tony}).
Comparing one month before and after the suspension, this channel also witnessed a drop of view counts by $-70\%$.

@@

As most YouTube channels upload videos on a weekly or even a monthly basis, a one-week suspension does not appear very restrictive at first sight. 
But we observed that the intervention was followed by a reduction in the number of view for the two channels investigated, with one channel even stopping its YouTube activity in the following months. 
Although more research is needed, it is possible that the one-week suspension following a strike might reduce the reach of active YouTube channels.

@@

@@

As we collected the recommendations of misinformation videos, we expected to find a high proportion of videos of questionable content.
But only {$18.4\%$} of the recommendations corresponded to $misleading$ or {\it hyper-partisan} videos.
This relatively small percentage hints towards that idea that YouTube is probably reducing the visibility of videos containing misinformation.
However, when we investigated misinformation at the channel level, we found that videos belonging to channels that are not highly authoritative were being often recommended.
It appears that the recommendation system of YouTube is reducing the visibility at the video level but not necessarily at the channel level.
